---
title: "RepData_Assignment1"
output: html_document
---

### Loading libraries
```{r}
library(plyr)
library(ggplot2)
library(lattice)
```

### Loading and preprocessing the data
```{r init}
data <- read.csv("activity.csv", header = T, sep = ',')
data.RemoveNa <- subset(data, is.na(data$steps) == F)
```

### What is mean total number of steps taken per day?
#### 1. Make a histogram of the total number of steps taken each day
```{r}
# Calculate the total number of steps per day
totalStepsPerDay <- ddply(
    data.RemoveNa, .(date), summarise, totalSteps=sum(steps))
# create the plot
barplot(totalStepsPerDay$totalSteps, names.arg = totalStepsPerDay$date,
        col = "black", main = "The total number of steps taken each day",
        xlab = "Date", ylab = "The total number of steps")
```

#### 2. Calculate the mean and median total number of steps taken per day
The mean number of steps taken per day
```{r}
mean(totalStepsPerDay$totalSteps)
```
The median number of steps taken per day
```{r}
median(totalStepsPerDay$totalSteps)
```

### What is the average daily activity pattern?
#### 1. Make a time series plot of the 5-minute interval (x-axis) and 
####    the average number of steps taken, averaged across all days (y-axis)
```{r}
# Calcluating the average number of steps taken in each 5-minite intervals
avg.PerInterval <- ddply(data.RemoveNa, .(interval), 
                         summarise, meanSteps = (mean(steps)))
# create the plot
plot(avg.PerInterval$interval, avg.PerInterval$meanSteps, axes = F, type = "l",
     col = "Black", 
     main = "The average number of steps taken in each 5-minite intervals",
     xlab = "Intervals",
     ylab = "The average number of steps taken")
axis(1, 
     at=c(0, 600, 1200, 1800, 2400), 
     label = c("0:00", "6:00", "12:00", "18:00", "24:00"))
axis(2)
```

#### 2. Which 5-minute interval, on average across all the days in the dataset,
#### contains the maximum number of steps?
```{r}
interval.maxMeanSteps <- avg.PerInterval[which.max(avg.PerInterval$meanSteps),]
interval <- interval.maxMeanSteps$interval
maxMeanSteps <- interval.maxMeanSteps$meanSteps
fromhour <- as.integer(interval/100)
frommin <- interval%%100
if (frommin >= 55){
    tohour <- fromhour + 1
    tomin <- (frommin + 5) %% 60
} else {
    tohour <- fromhour
    tomin <- frommin + 5
}
```
So the interval that contains the maximum number of step is 
`r fromhour`:`r frommin` -- `r tohour`:`r tomin`, 
and the average steps is `r maxMeanSteps`.

### Imputing missing values
#### 1. Calculate and report the total number of missing values in the dataset
```{r}
sum(is.na(data$steps))
```

#### 2. Devise a strategy for filling in all of the missing values
#### 3. Create a new dataset that is equal to the original dataset 
#### but with the missing data filled in
```{r}
data.impute <- data
for (i in 1:nrow(data.impute)){
    if (is.na(data.impute$steps[i]))
        data.impute$steps[i] <- 
            avg.PerInterval$meanSteps[
                which(data.impute$interval[i] == avg.PerInterval$interval)]
}
head(data.impute)
```

#### 4. Make a histogram of the total number of steps taken each day and 
#### Calculate the mean and median total number of steps taken per day. 
#### Do these values differ from the estimates from the first part of the 
#### assignment? What is the impact of imputing missing data on the estimates 
#### of the total daily number of steps?
```{r}
# Calculate the total number of steps per day for the imputed data
totalStepsPerDay.impute <- ddply(data.impute, .(date),
                                   summarise, totalSteps = sum(steps))

# create the plot for the imputed data
barplot(totalStepsPerDay.impute$totalSteps, 
        names.arg = totalStepsPerDay.impute$date,
        col = "black", 
        main = "The total number of steps taken each day for the imputed data",
        xlab = "Date", ylab = "The total number of steps")
```

Calculate and report the mean and median total number of steps taken 
per day for the imputed data
```{r}
mean(totalStepsPerDay.impute$totalSteps)
median(totalStepsPerDay.impute$totalSteps)
```

Differences between the values for the imputed data and 
the ones for the origin ones
```{r}
mean(totalStepsPerDay.impute$totalSteps) - mean(totalStepsPerDay$totalSteps)
median(totalStepsPerDay.impute$totalSteps) - median(totalStepsPerDay$totalSteps)
sum(totalStepsPerDay.impute$totalSteps) - sum(totalStepsPerDay$to)
```
So the mean didn't change after the imputing, and the median only 
slightly changed,  but the total steps increased.

### Are there differences in activity patterns between weekdays and weekends?
#### 1. Create a new factor variable in the dataset with two levels -- 
#### "weekday" and "weekend" indicating whether a given date is a weekday or
#### weekend day.
```{r}
Sys.setlocale("LC_TIME", "Chinese") 
dates <- weekdays(as.Date(data.impute$date))
data.impute$dayType <- 
    ifelse(dates %in% 
               c("Saturday", "Sunday"), "weekend", "weekday")
```

#### 2.Make a panel plot containing a time series plot of the 5-minute interval
#### and the average number of steps taken, averaged across all weekday days or
#### weekend days. 
```{r}
# calcluate the average for each interval
avgSteps.byDayType <- ddply(data.impute, .(interval, dayType), 
                            summarise, meanSteps = mean(steps))

# create the plot
xyplot(meanSteps ~ interval | dayType, 
       data = avgSteps.byDayType, 
       layout = c(1, 2), 
       type = "l", 
       xlab = "Interval", 
       ylab = "The average number of steps")

# create the plot using ggplot
g <- ggplot(avgSteps.byDayType, aes(interval, meanSteps))
g + facet_grid(dayType~.) + 
    labs(x = "Interval") + 
    labs(y = "The average number of steps") +
    geom_line(aes(color = dayType))
```
